\mytitle

### 2015-08-01 MSc_diss.tex
Started subsection about ANs formation

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common. Its photolysis and further oxidation contributes to radical generation:
\begin{equation}\label{reac:HCHO+hv=H+HCO}
HCHO + hv \rightarrow H + HCO
\end{equation}
\begin{equation}\label{reac:HCO+O2=HO2+CO}
HCO + O_2 \rightarrow HO_2 + CO
\end{equation}
\begin{equation}\label{reac:H+O2=HO2}
!!!sameH + O_2 \xrightarrow{M} HO_2
\end{equation}
therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section. Their role is described in other sections as fas as it is necessary.

\subsection{Alkyl nitrates}
Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}. This mechanism is present in the following section.

@@

@@

\subsubsection*{Bertman}
Below is the simplified scheme of atmospheric hydrocarbon oxidation leading to the formation of alkyl nitrates (see also Fig. \ref{fig:scheme_bertman}). It begins from the process of $H$ abstraction by the $OH$ radical where the fraction of the primary or secondary radical is denoted by $\alpha_1$.
\begin{equation} \label{eq:alkane_oh}
RH + \bullet OH \xrightarrow{k_1, \alpha_1} R\bullet + H_2O
\end{equation}
The resulting radical rapidly reacts with $O_2$ and forms an alkyl peroxy radical ($R\bullet$):
\begin{equation} \label{eq:rad_o2}
R\bullet + O_2 \xrightarrow{k_2} ROO\bullet
\end{equation}
Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate. The former reaction is a crucial in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
ROO\bullet + NO &\xrightarrow{k_3, (1-\alpha_3)} RO\bullet + NO_2 \label{eq:peroxy_no1}\\
ROO\bullet + NO &\xrightarrow{k_3, \alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of nitrate products via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

@@

Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
\begin{subequations} \label{eq:an_form_loss0}
\begin{align}
RH &\xrightarrow{\beta k_A} RONO2 \label{eq:an_form_loss1}\\
RONO_2 &\xrightarrow{k_B} \mathit{products} \label{eq:an_form_loss2}
\end{align}
\end{subequations}
where $k_A = k_1[OH]$ and $k_B=j_4 + k_5[OH]$. The factor $\beta=\alpha_1\alpha_3$ takes into account the fraction of $H$ atom abstraction from the particular hydrocarbon (with the reaction rate $k_1$) and the branching ratios leading to the nitrate. The photolytic loss rate of $RONO_2$ is denoted by $j_4$, and $k_5$ is the rate of alkyl nitrate destruction by $OH$.

@@

A further good approximation is that there is no alkyl nitrate initially (\textit{i.e.} no direct emissions), hence, the second term in eq. (\eqref{eq:integral}) becomes zero. Finally, the evolution of alkyl nitrates in an air mass can be modelled using the following formula.
\begin{equation} \label{eq:model}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)
\end{equation}
It has to be noted that eq. \eqref{eq:model} assumes no mixing with surrounding air. This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs. \eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}.

>>>>

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common. Its photolysis and further oxidation contributes to radical generation:
\begin{equation}\label{reac:HCHO+hv=H+HCO}
HCHO + hv \rightarrow H + HCO
\end{equation}
\begin{equation}\label{reac:HCO+O2=HO2+CO}
HCO + O_2 \rightarrow HO_2 + CO
\end{equation}
\begin{equation}\label{reac:H+O2=HO2}
!!!sameH + O_2 \xrightarrow{M} HO_2
\end{equation}
therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section. Their role is described in other sections as far as it is necessary.

\subsection{Alkyl nitrates}
Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}. This mechanism is present in the following section.

@@

The photochemical formation of the alkyl nitrates starts when the chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration. In this case the reaction between of the alkyl peroxy radicals ($RO_2$ where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way:
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{eq:peroxy_no1}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
The former reaction is a crucial since it can serve as a termination step

in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.

In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

@@

\subsubsection*{Bertman}
Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate. The former reaction is a crucial in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
ROO\bullet + NO &\xrightarrow{k_3, (1-\alpha_3)} RO\bullet + NO_2 \label{eq:peroxy_no1}\\
ROO\bullet + NO &\xrightarrow{k_3, \alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

@@

[Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
\begin{subequations} \label{eq:an_form_loss0}
\begin{align}
RH &\xrightarrow{\beta k_A} RONO2 \label{eq:an_form_loss1}\\
RONO_2 &\xrightarrow{k_B} \mathit{products} \label{eq:an_form_loss2}
\end{align}
\end{subequations}
where $k_A = k_1[OH]$ and $k_B=j_4 + k_5[OH]$. The factor $\beta=\alpha_1\alpha_3$ takes into account the fraction of $H$ atom abstraction from the particular hydrocarbon (with the reaction rate $k_1$) and the branching ratios leading to the nitrate. The photolytic loss rate of $RONO_2$ is denoted by $j_4$, and $k_5$ is the rate of alkyl nitrate destruction by $OH$.

@@

A further good approximation is that there is no alkyl nitrate initially (\textit{i.e.} no direct emissions), hence, the second term in eq. (\eqref{eq:integral}) becomes zero. Finally, the evolution of alkyl nitrates in an air mass can be modelled using the following formula.
\begin{equation} \label{eq:model}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)
\end{equation}
It has to be noted that eq. \eqref{eq:model} assumes no mixing with surrounding air. This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs. \eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}].
